Rheologic constraints on the upper mantle from five years of postseismic deformation following the El Mayor-Cucapah earthquake

### 2016-06-29 JGRRevision/ElMayor.tex
added supplementary material

\begin{abstract}
Five years of postseismic deformation following the  Mw=7.2 El Mayor-Cucapah earthquake reveals transient deformation that decays back to its pre-earthquake trend after about three years at epicentral distances greater than ${\sim}200$ km. At shorter distances, the rapid transience decays to a sustained rate which exceeds its pre-earthquake trend.  We attempt to determine the mechanisms driving this deformation, where we consider afterslip at seismogenic depths and viscoelastic relaxation in the lower crust and upper mantle as candidate mechanisms.  We find that early, rapid, near-field deformation can be explained with afterslip on the fault that ruptured coseismically, while the later, sustained, near-field deformation can be explained with either continued afterslip in an effectively elastic lower crust, or lower crustal viscoelastic relaxation with a steady-state viscosity of ${\sim}10^{19}$ Pa s. The trend in far-field deformation is best explained with a transient viscosity of ${\sim}10^{18}$ Pa s in the upper mantle. We argue that a transient rheology in the mantle is preferable over a Maxwell rheology because it better predicts the decay in postseismic deformation, and also because it does not conflict with the generally higher, steady-state viscosities inferred from studies of geophysical processes occurring over longer time-scales.
\end{abstract}

\section{Introduction}
Ground deformation in the years following a large ($\gtrsim$Mw7) earthquake can be used to gain insight into the mechanical behavior of the crust and upper mantle. The interpretations of postseismic deformation are not always conclusive because multiple postseismic deformation mechanisms, such as afterslip or viscoelastic relaxation in the lower crust and upper mantle, can have qualitatively similar surface expressions \citep[e.g.][]{Savage1990}. This complication can potentially be remedied by having a sufficient number of well positioned observations of the postseismic displacement field \citep{Hearn2003}.  Owing to the dense geodetic network deployed throughout the 2000s as part of the Plate Boundary Observatory, the postseismic deformation following the April 4, 2010, Mw7.2 El Mayor-Cucapah earthquake in Baja California was observed at more GPS stations than any other earthquake in California to date. With such a large collection of data, we attempt to discern the mechanisms driving postseismic deformation following the El Mayor-Cucapah earthquake.

@@

\begin{figure}
\includegraphics[scale=0.8]{Figures/ContextMap} 
\caption{Map of the region considered in this study.  The large focal mechanism is for the El Mayor-Cucapah earthquake, and the three small focal mechanisms are for the Ocotillo earthquake and the two main shocks during the Brawley swarm.  The black dots indicate the locations of GPS stations used in this study.  The fault geometry used in this study is shown in cyan where dashed lines indicate buried edges of the fault segments.  The green and red boxes demarcate the extent of the near-field and far-field maps (Figures \ref{fig:NearField} and \ref{fig:FarField}).  Stations inside the blue sector, which highlights the area within 10$^\circ$ of the El Mayor-Cucapah P axis, are used in Figures \ref{fig:RecordSectionElastic}, \ref{fig:RecordSection1}, and \ref{fig:RecordSection2}}.       
\label{fig:ContextMap}
\end{figure}

\section{Data Processing}\label{sec:Data}
We use continuous GPS position time series provided by University Navstar Consortium (UNAVCO) for stations within a 400 km radius about the El Mayor-Cucapah epicenter. We collectively describe the coseismic and postseismic displacements resulting from the El Mayor-Cucapah earthquake as $u_\mathrm{post}(t)$.  We consider the GPS position time series $u_\mathrm{obs}(t)$ to be the combination of $u_\mathrm{post}(t)$, secular tectonic deformation, annual and semi-annual oscillations, and coseismic offsets from significant earthquakes over the time span of this study.  The June 14, 2010, Mw5.8 Ocotillo earthquake and the Brawley swarm, which included an Mw5.5 and an Mw5.4 event on August 26, 2012, (Figure \ref{fig:ContextMap}) are the only earthquakes that produced noticeable displacements in any of the time series.  We treat the displacements resulting from the Brawley swarm as a single event because the daily solutions provided by UNAVCO cannot resolve the separate events.  Although the Ocotillo earthquake had its own series of aftershocks \citep{Hauksson2011}, neither the Ocotillo earthquake nor the Brawley swarm produced detectable postseismic deformation.  We model displacements resulting from these events with only a Heaviside function, $H(t)$, describing the coseismic offsets.  We then model $u_\mathrm{obs}(t)$ as

@@

The October 16, 1999, Mw7.1 Hector Mine earthquake, which occurred ${\sim}270$ km north of the El Mayor-Cucapah epicenter, produced transient postseismic deformation which we do not wish to model, either mechanically or through empirical line fitting.  We thus restrict our analysis to deformation observed six years after the Hector Mine earthquake, which is when postseismic velocities at sites proximal to the Hector Mine epicenter are approximately constant \citep{Savage2009}. When appraising our model fit in Section \ref{sec:Model}, we see some systematic residuals in the vicinity of the Hector Mine epicenter, which may be the result of errors in the assumption that the trend in Hector Mine postseismic deformation is linear after six years.

@@

\begin{figure}
\includegraphics[scale=0.8]{Figures/FilterP496}
\centering
\caption{Left panels show GPS time series from UNAVCO (black) and the predicted displacement (blue) from eq. (\ref{TimeSeriesModel}) for a near-field station.  Red lines indicate the times of the El Mayor-Cucapah and Ocotillo earthquake. The right panels show estimated coseismic and postseismic displacements $u_\mathrm{post}$ which are extracted from the predicted displacements.  The 68\% confidence interval is shown in light blue.}
\label{fig:P496}
\end{figure}

\begin{figure}
\includegraphics[scale=0.8]{Figures/FilterP619}
\centering
\caption{same as Figure \ref{fig:P496} but for a far-field station.} 
\label{fig:P619}
\end{figure}

@@

\begin{figure}
\includegraphics[scale=0.8]{Figures/MapViewNearField}
\centering 
\caption{Near-field coseismic and cumulative postseismic displacements over the indicated time periods (black) as well as predicted displacements for our preferred model from Section \ref{sec:FullInversion} (green).  Observed vertical deformation is shown as an interpolated field and predicted vertical displacements are shown within the circles at the base of each vector.  Note that the interpolant is not well constrained in Mexico where there is no data available.}
\label{fig:NearField}
\end{figure}

\begin{figure}
\includegraphics[scale=0.8]{Figures/MapViewFarField}
\centering 
\caption{Same as Figure \ref{fig:NearField} but for far-field stations.}
\label{fig:FarField}
\end{figure}

@@

Our coseismic slip and afterslip solutions are shown in Figure \ref{fig:ElasticSlip}.  As with \citet{Rollins2015}, we find that a large amount of afterslip on the southern fault segment is required to explain the observations. The potency of our inferred coseismic slip is $3.2\times10^9\ \mathrm{m}^3$, equivalent to a Mw7.28 earthquake when assuming a shear modulus of 32 GPa.  The potency of our inferred cumulative five years of afterslip is $6.1\times10^9\ \mathrm{m}^3$, equivalent to a Mw7.46 earthquake, which is unrealistically large if we consider afterslip to be driven by coseismically induced stresses.  Figure \ref{fig:RecordSectionElastic} shows the time series for the observed and predicted postseismic displacements at stations along the El Mayor-Cucapah P axis.  We show the radial component of displacements with respect to the El Mayor-Cucapah epicenter and we also rescale the displacements so that the difference between the minimum and maximum observed displacements are the same for each station.  Our elastic slip model accurately describes near-field postseismic deformation and systematically underestimates postseismic deformation at epicentral distances ${\gtrsim}150$ km.  When the fault segments used in the inversion are extended down to 30 km depth, rather than 15 km, the systematic far-field residuals are smaller but remain apparent. Because an elastic model requires an unrealistic amount of afterslip and is unable to predict far-field deformation, we move on to consider viscoelastic models in the next section.

\begin{figure}
\includegraphics[scale=0.8]{Figures/ElasticSlip}
\caption{Coseismic slip and cumulative afterslip over the indicated intervals for the elastic model from Section \ref{sec:ElasticInversion}.  Color indicates the magnitude of slip while arrows indicate the motion of the hanging wall.}
\label{fig:ElasticSlip}
\end{figure}

\begin{figure}
\includegraphics[scale=0.8]{Figures/RecordSectionElastic}
\caption{Scaled radial component of observed postseismic displacements $u_\mathrm{obs}$ (black) and displacements predicted by the best fitting elastic model $u_\mathrm{pred}$ (blue).  Downward motion indicates that the station is moving toward the El Mayor-Cucapah epicenter.  Displacement time series are scaled so that the minimum and maximum observed values lie on the grid lines.  The 68\% confidence interval for the observed displacements are shown in gray.}
\label{fig:RecordSectionElastic}
\end{figure}

\subsection{Early Postseismic Inversion}\label{sec:InitialInversion}

For any linear viscoelastic rheology of the crust and mantle, postseismic displacements resulting from time dependent fault slip can be described as

@@

\begin{figure}
\includegraphics[scale=0.8]{Figures/rheology}
\centering 
\caption{Schematic illustration of the rheologic models considered in this paper as well as their effective viscosities.}
\label{fig:Rheology}
\end{figure}

@@

\begin{figure}
\includegraphics[scale=0.8]{Figures/Cancellation}
\caption{Displacements resulting from fault slip at lower crustal depths (a), and initial velocities resulting from subsequent relaxation of a viscoelastic lower crust (b).  The fault segment dips $75^\circ$ to the north-east and its surface projection is outlined in magenta.  The highlighted area on the fault extends from 15 to 30 km depth and indicates where 1 meter of right-lateral slip was imposed.  The elastic properties of the crust and mantle are the same as in Table \ref{tab:MaterialProperties}, and $\eta_\mathrm{eff}$  is $10^{18}$ Pa s in the lower crust.  Vertical displacements are interpolated between station locations.}
\label{fig:LowerCrust}
\end{figure}

@@

\begin{figure}
\includegraphics[scale=0.8]{Figures/RecordSectionDuration}
\centering 
\caption{Observed postseismic displacements (black) and best fitting predictions of eq. (\ref{ApproxForward}) to 5.0 (blue), 3.0 (green), and 0.8 (yellow) years of the postseismic data.} 
\label{fig:RecordSection1}
\end{figure}

@@

\begin{figure}
\includegraphics[scale=0.8]{Figures/EffectiveViscosity}
\centering 
\caption{Effective viscosities and associated fluidities inferred by fitting eq. (\ref{ApproxForward}) to the first 0.8 years of postseismic displacements. 95\% confidence intervals, estimated from bootstrapping, are indicated by color fields.}
\label{fig:EffectiveViscosity}
\end{figure}

@@

\begin{figure}
\includegraphics[scale=0.8]{Figures/InitialSlip}
\caption{Coseismic slip and afterslip inferred by fitting eq. (\ref{ApproxForward}) to the first 0.8 years of postseismic displacements.}
\label{fig:InitialSlip}
\end{figure}

@@

We first assume that the crust and mantle can be described with a Maxwell rheology, and we set the steady-state viscosity $\eta_\mathrm{M}$ equal to our inference of $\eta_{\mathrm{eff}}$.  We compute $f$ and $g$ from eq. (\ref{GeneralForward}) using Pylith, and we use the same spatial and temporal discretization of $s$ as in Sections \ref{sec:ElasticInversion} and \ref{sec:InitialInversion}. We estimate $s$ using linear least squares and find a misfit of $\bar\chi^2=37.4$. For comparison, $\bar\chi^2=35.3$ for the elastic model from Section \ref{sec:ElasticInversion}.  The Maxwell viscoelastic model has a larger misfit because it tends to overestimate the rate of deformation after about three years (Figure \ref{fig:RecordSection2}). Since our initial estimates of $\eta_\mathrm{eff}$ may be biased towards overestimating viscosities, we have also performed the slip inversion where we use uniformly lower viscosities in the crust and mantle. However, decreasing the viscosity only increases the misfit.  Although, the viscosities used here are consistent with the successful Maxwell viscoelastic models found by \citet{Rollins2015} and \citet{Spinler2015}, which had mantle viscosities on the order of $10^{18}$ Pa s and relatively higher lower crustal viscosities, we find that such a model is incapable of describing the entire postseismic time series.  \citet{Pollitz2001} similarly recognized this deficiency in a Maxwell rheology, which then motivated their exploration of a Burgers rheology upper mantle \citep{Pollitz2003}.

Instead of exploring a Burgers rheology mantle, which introduces two new parameters that need to be estimated, the transient viscosity $\eta_{K}$ and transient shear modulus $\mu_{K}$, we first consider a Zener rheology for the mantle, which only introduces one additional unknown parameter, $\mu_{K}$.  We assume that the lower crust still has a Maxwell rheology. The steady-state viscosity in the crust and the transient viscosity in the mantle are set equal to the inferred effective viscosities.  We then estimate the ratio of shear moduli $\frac{\mu_K}{\mu}$. We compute nine different sets of Green's functions, $f$ and $g$, where we assume values of $\frac{\mu_K}{\mu}$ ranging from 0 to 1. The former being a degenerate case where the Zener model reduces to the above Maxwell model.  We estimate coseismic slip and afterslip for each realization of $\frac{\mu_K}{\mu}$.  The shear moduli ratio that yields the best prediction to the observed postseismic displacements is found to be $0.375$ which produces a misfit of $\bar\chi^2=31.2$ (Figure \ref{fig:ShearModulusRatio}).  The improvement in the Zener model over the Maxwell model can be clearly seen in the fit to the far-field data (Figure \ref{fig:RecordSection2}). The Zener model does a significantly better job at explaining the transient rate of far-field deformation throughout the five years.

\begin{figure}
\includegraphics[scale=0.8]{Figures/RatioMisfit}
\centering 
\caption{Mean chi-squared value for the best fitting slip model as a function of the transient shear modulus relative to the elastic shear modulus in a Zener upper mantle. Large dot indicates our preferred ratio.}
\label{fig:ShearModulusRatio}
\end{figure}

\begin{figure}
\includegraphics[scale=0.8]{Figures/RecordSectionFinal}
\centering 
\caption{Observed postseismic displacements (black). Predicted postseismic displacements for the best fitting slip model when assuming a Maxwell viscoelastic lower crust and upper mantle (blue).  Predicted postseismic displacements for the best fitting slip model when assuming a Maxwell viscoelastic lower crust and a Zener viscoelastic upper mantle (green).  The effective viscosities are the same for both models and are shown in Figure \ref{fig:EffectiveViscosity}.}
\label{fig:RecordSection2}
\end{figure}

@@

The regularized inference of coseismic slip and afterslip for our preferred Zener model is shown in Figure \ref{fig:FinalSlip}.  The inferred coseismic potency is $3.0\times10^{9}\ \mathrm{m}^3$, equivalent to a Mw7.26 earthquake, and most of the slip is shallow and on the Sierra Cucapah fault segment.  The potency of five years of afterslip is $1.1\times10^{9}\ \mathrm{m}^3$. Most of the afterslip in our preferred model occurs within the first year after the earthquake and coincides with the location of our inferred coseismic slip. Inferred afterslip within the first year is accounting for the most rapid near-field transient deformation. After one year, afterslip is inferred to be deeper down on the Sierra Cucapah segment, which is describing much of the sustained near-field postseismic deformation.  We emphasize, that the GPS station closest to where we infer afterslip, P496, is still about 30 km away, which is too far for us to conclusively argue for sustained localized deformation rather than shallow distributed deformation.  The deep afterslip inferred after one year could potentially be describing deformation resulting from lower crustal flow. To test this, we have modified our preferred model by decreasing the lower crustal viscosity from $5.91\times10^{19}$ Pa s to $1\times10^{19}$ Pa s, which is still consistent with our viscosity inference from Section \ref{sec:InitialInversion}, and inverted for fault slip.  We find that a model with a weaker lower crust adequately describes the postseismic displacements without any afterslip after one year, while still requiring about the same amount of afterslip over the first year. We do believe that the early, shallow afterslip on the Sierra Cucapah segment is a robust feature in our preferred model, while we are not confident in our inference of later deep afterslip.

the postseismic displacements predicted by our preferred model are shown in Figures \ref{fig:NearField}, \ref{fig:FarField} and \ref{fig:RecordSection2}.  Overall, the trends in the near-field and far-field transient deformation are accurately described.  In particular, the trends in far-field deformation are much better described by our preferred Zener model than either an elastic model or a model with a Maxwell viscoelastic mantle (Figure \ref{fig:RecordSection2}).  There are a few areas where we have notable misfit.  Most of our misfit is for the near-field stations in the Imperial Valley, and we attribute this misfit to our relatively simple fault geometry, which does not account for potential fault slip in the Imperial Valley triggered by the El Mayor-Cucapah earthquake \citep{Wei2011a,Wei2015}. In particular, we are unable to model the sustained, rapid rate of deformation at station P496, which suggests that this station could be influenced by a more localized deformation mechanism than is considered in this study.  Additionally, we see systematic misfit in the later postseismic period west of the Landers and Hector Mine earthquakes, which may be the result of unmodeled postseismic deformation resulting from those earthquakes.  Lastly, there are clear discrepancies between the observed and predicted vertical deformation following the first year after the El Mayor Cucapah earthquake. We observe a broad uplift throughout Southern California, which is inconsistent with any postseismic model.

\begin{figure}
\includegraphics[scale=0.8]{Figures/FinalSlip}
\centering 
\caption{Inferred coseismic slip and afterslip when assuming a Maxwell rheology in the lower crust and a Zener rheology in the upper mantle.  The transient viscosity $\eta_K$ in the mantle and steady-state viscosity $\eta_M$ in the crust are set equal to the effective viscosities from Figure \ref{fig:EffectiveViscosity}. We set $\frac{\mu_K}{\mu}=0.375$ in the upper mantle.}
\label{fig:FinalSlip}
\end{figure}

@@

An additional deficiency with the Maxwell rheology is that it predicts a steady decay in the rate of postseismic deformation over time, which fails to describe the commonly observed rapid, early transience followed by a relatively steady rate of postseismic deformation.  One could explain the early transient postseismic deformation with fault creep and the later phase with relaxation in a Maxwell viscoelastic lower crust and upper mantle \citep[e.g][]{Hearn2009,Johnson2009}. However, postseismic deformation at distances greater than ${\sim}200$ km from the El Mayor-Cucapah epicenter can only be attributed to viscoelastic relaxation \citep{Freed2007a} and we have demonstrated that the far-field deformation cannot be explained with a Maxwell rheology (Figure \ref{fig:RecordSection2}).

>>>>

\begin{abstract}

Five years of postseismic deformation following the  Mw=7.2 El Mayor-Cucapah earthquake reveals transient deformation that decays back to its pre-earthquake trend after about three years at epicentral distances greater than ${\sim}200$ km. At shorter distances, the rapid transience decays to a sustained rate which exceeds its pre-earthquake trend.  We attempt to determine the mechanisms driving this deformation, where we consider afterslip at seismogenic depths and viscoelastic relaxation in the lower crust and upper mantle as candidate mechanisms.  We find that early, rapid, near-field deformation can be explained with afterslip on the fault that ruptured coseismically, while the later, sustained, near-field deformation can be explained with either continued afterslip in an effectively elastic lower crust, or lower crustal viscoelastic relaxation with a steady-state viscosity of ${\sim}10^{19}$ Pa s. The trend in far-field deformation is best explained with a transient viscosity of ${\sim}10^{18}$ Pa s in the upper mantle. We argue that a transient rheology in the mantle is preferable over a Maxwell rheology because it better predicts the decay in postseismic deformation, and also because it does not conflict with the generally higher, steady-state viscosities inferred from studies of geophysical processes occurring over longer time-scales.
\end{abstract}

\section{Introduction}
Ground deformation in the years following a large (Mw$\gtrsim$7) earthquake can be used to gain insight into the mechanical behavior of the crust and upper mantle. The interpretations of postseismic deformation are not always conclusive because multiple postseismic deformation mechanisms, such as afterslip or viscoelastic relaxation in the lower crust and upper mantle, can have qualitatively similar surface expressions \citep[e.g.][]{Savage1990}. This complication can potentially be remedied by having a sufficient number of well positioned observations of the postseismic displacement field \citep{Hearn2003}.  Owing to the dense geodetic network deployed throughout the 2000s as part of the Plate Boundary Observatory, the postseismic deformation following the April 4, 2010, Mw=7.2 El Mayor-Cucapah earthquake in Baja California was observed at more GPS stations than any other earthquake in California to date. With such a large collection of data, we attempt to discern the mechanisms driving postseismic deformation following the El Mayor-Cucapah earthquake.

@@

\begin{figure}
\includegraphics[scale=0.8]{Figures/ContextMap} 
\caption{Map of the region considered in this study.  The large focal mechanism is for the El Mayor-Cucapah earthquake, and the three small focal mechanisms are for the Ocotillo earthquake and the two main shocks during the Brawley swarm.  The black dots indicate the locations of GPS stations used in this study.  The fault geometry used in this study is shown in cyan where dashed lines indicate buried edges of the fault segments.  The green and red boxes demarcate the extent of the near-field and far-field maps (Figures \ref{fig:NearField} and \ref{fig:FarField}).  Stations inside the blue sector, which highlights the area within 10$^\circ$ of the El Mayor-Cucapah P axis, are used in Figures \ref{fig:RecordSectionMain} and \ref{fig:RecordSection1}.}       
\label{fig:ContextMap}
\end{figure}

\section{Data Processing}\label{sec:Data}
We use continuous GPS position time series provided by University Navstar Consortium (UNAVCO) for stations within a 400 km radius about the El Mayor-Cucapah epicenter. We collectively describe the coseismic and postseismic displacements resulting from the El Mayor-Cucapah earthquake as $u_\mathrm{post}(t)$.  We consider the GPS position time series $u_\mathrm{obs}(t)$ to be the combination of $u_\mathrm{post}(t)$, secular tectonic deformation, annual and semi-annual oscillations, and coseismic offsets from significant earthquakes over the time span of this study.  The June 14, 2010, Mw=5.8 Ocotillo earthquake and the Brawley swarm, which included an Mw=5.5 and an Mw=5.4 event on August 26, 2012, (Figure \ref{fig:ContextMap}) are the only earthquakes that produced noticeable displacements in any of the time series.  We treat the displacements resulting from the Brawley swarm as a single event because the daily solutions provided by UNAVCO cannot resolve the separate events.  Although the Ocotillo earthquake had its own series of aftershocks \citep{Hauksson2011}, neither the Ocotillo earthquake nor the Brawley swarm produced detectable postseismic deformation.  We model displacements resulting from these events with only a Heaviside function, $H(t)$, describing the coseismic offsets.  We then model $u_\mathrm{obs}(t)$ as

@@

The October 16, 1999, Mw=7.1 Hector Mine earthquake, which occurred ${\sim}270$ km north of the El Mayor-Cucapah epicenter, produced transient postseismic deformation which we do not wish to model, either mechanically or through empirical line fitting.  We thus restrict our analysis to deformation observed six years after the Hector Mine earthquake, which is when postseismic velocities at sites proximal to the Hector Mine epicenter are approximately constant \citep{Savage2009}. When appraising our model fit in Section \ref{sec:Model}, we see some systematic residuals in the vicinity of the Hector Mine epicenter, which may be the result of errors in the assumption that the trend in Hector Mine postseismic deformation is linear after six years.

@@

\begin{figure}
\includegraphics[scale=0.9]{Figures/FilterP496}
\centering
\caption{Left panels show GPS time series from UNAVCO (black) and the predicted displacement (blue) from eq. (\ref{TimeSeriesModel}) for a near-field station.  Red lines indicate the times of the El Mayor-Cucapah and Ocotillo earthquake. The right panels show estimated coseismic and postseismic displacements $u_\mathrm{post}$ which are extracted from the predicted displacements.  The 68\% confidence interval is shown in light blue.}
\label{fig:P496}
\end{figure}

\begin{figure}
\noindent\includegraphics[scale=0.9]{Figures/FilterP619}
\centering
\caption{same as Figure \ref{fig:P496} but for a far-field station.} 
\label{fig:P619}
\end{figure}

@@

\begin{figure}
\includegraphics[scale=0.9]{Figures/MapViewNearField}
\centering 
\caption{Near-field coseismic and cumulative postseismic displacements over the indicated time periods (black) as well as predicted displacements for our preferred model from Section \ref{sec:FullInversion} (green).  Observed vertical deformation is shown as an interpolated field and predicted vertical displacements are shown within the circles at the base of each vector.  Note that the interpolant is not well constrained in Mexico where there is no data available.}
\label{fig:NearField}
\end{figure}

\begin{figure}
\includegraphics[scale=0.9]{Figures/MapViewFarField}
\centering 
\caption{Same as Figure \ref{fig:NearField} but for far-field stations.}
\label{fig:FarField}
\end{figure}

@@

Our coseismic slip and afterslip solutions are shown in Figure \ref{fig:ElasticSlip}.  As with \citet{Rollins2015}, we find that a large amount of afterslip on the southern fault segment is required to explain the observations. The potency of our inferred coseismic slip is $3.2\times10^9\ \mathrm{m}^3$, equivalent to a Mw=7.28 earthquake when assuming a shear modulus of 32 GPa.  The potency of our inferred cumulative five years of afterslip is $6.1\times10^9\ \mathrm{m}^3$, equivalent to a Mw=7.46 earthquake, which is unrealistically large if we consider afterslip to be driven by coseismically induced stresses.  Figure \ref{fig:RecordSectionMain} shows the time series for the observed and predicted postseismic displacements at stations along the El Mayor-Cucapah P axis.  We show the radial component of displacements with respect to the El Mayor-Cucapah epicenter and we also rescale the displacements so that the difference between the minimum and maximum observed displacements are the same for each station.  Our elastic slip model accurately describes near-field postseismic deformation and systematically underestimates postseismic deformation at epicentral distances ${\gtrsim}150$ km.  When the fault segments used in the inversion are extended down to 30 km depth, rather than 15 km, the systematic far-field residuals are smaller but remain apparent. Because an elastic model requires an unrealistic amount of afterslip and is unable to predict far-field deformation, we move on to consider viscoelastic models in the next section.

\begin{figure}
\includegraphics[scale=0.9]{Figures/ElasticSlip}
\caption{Coseismic slip and cumulative afterslip over the indicated intervals for the elastic model from Section \ref{sec:ElasticInversion}.  Color indicates the magnitude of slip while arrows indicate the motion of the hanging wall.}
\label{fig:ElasticSlip}
\end{figure}

\begin{figure}
\includegraphics[scale=0.9]{Figures/RecordSectionFinal}
\caption{Scaled radial component of postseismic displacements. Downward motion indicates that the station is moving toward the El Mayor-Cucapah epicenter.  Displacement time series are scaled so that the minimum and maximum observed values lie on the grid lines.  The observed postseismic displacements, $u_\mathrm{obs}$ are shown in black with gray indicating the 68\% confidence interval.  The displacements predicted by the best fitting elastic model $u_\mathrm{pred}$ are shown in red.  The blue and green lines are the predicted postseismic displacements for the models discussed in Section 3.3. The blue lines show the predicted displacements for the model with a Maxwell viscoelastic lower crust and upper mantle.  The green line shows the predicted displacements for the model with a Maxwell viscoelastic lower crust and a Zener viscoelastic upper mantle.  The effective viscosities are the same for both models and are shown in Figure \ref{fig:EffectiveViscosity}.}
\label{fig:RecordSectionMain}
\end{figure}

\subsection{Early Postseismic Inversion}\label{sec:InitialInversion}
For any linear viscoelastic rheology of the crust and mantle, postseismic displacements resulting from time dependent fault slip can be described as

@@

\begin{figure}
\includegraphics[scale=0.9]{Figures/rheology}
\centering 
\caption{Schematic illustration of the rheologic models considered in this paper as well as their effective viscosities.}
\label{fig:Rheology}
\end{figure}

@@

\begin{figure}
\includegraphics[scale=0.9]{Figures/Cancellation}
\caption{Displacements resulting from fault slip at lower crustal depths (a), and initial velocities resulting from subsequent relaxation of a viscoelastic lower crust (b).  The fault segment dips $75^\circ$ to the north-east and its surface projection is outlined in magenta.  The highlighted area on the fault extends from 15 to 30 km depth and indicates where 1 meter of right-lateral slip was imposed.  The elastic properties of the crust and mantle are the same as in Table \ref{tab:MaterialProperties}, and $\eta_\mathrm{eff}$  is $10^{18}$ Pa s in the lower crust.  Vertical displacements are interpolated between station locations.}
\label{fig:LowerCrust}
\end{figure}

@@

\begin{figure}
\includegraphics[scale=0.9]{Figures/RecordSectionDuration}
\centering 
\caption{Observed postseismic displacements (black) and best fitting predictions of eq. (\ref{ApproxForward}) to 5.0 (blue), 3.0 (green), and 0.8 (yellow) years of the postseismic data.} 
\label{fig:RecordSection1}
\end{figure}

@@

\begin{figure}
\includegraphics[scale=0.9]{Figures/EffectiveViscosity}
\centering 
\caption{Effective viscosities and associated fluidities inferred by fitting eq. (\ref{ApproxForward}) to the first 0.8 years of postseismic displacements. 95\% confidence intervals, estimated from bootstrapping, are indicated by color fields.}
\label{fig:EffectiveViscosity}
\end{figure}

@@

\begin{figure}
\includegraphics[scale=0.9]{Figures/InitialSlip}
\caption{Coseismic slip and afterslip inferred by fitting eq. (\ref{ApproxForward}) to the first 0.8 years of postseismic displacements.}
\label{fig:InitialSlip}
\end{figure}

@@

We first assume that the crust and mantle can be described with a Maxwell rheology, and we set the steady-state viscosity $\eta_\mathrm{M}$ equal to our inference of $\eta_{\mathrm{eff}}$.  We compute $f$ and $g$ from eq. (\ref{GeneralForward}) using Pylith, and we use the same spatial and temporal discretization of $s$ as in Sections \ref{sec:ElasticInversion} and \ref{sec:InitialInversion}. We estimate $s$ using linear least squares and find a misfit of $\bar\chi^2=37.4$. For comparison, $\bar\chi^2=35.3$ for the elastic model from Section \ref{sec:ElasticInversion}.  The Maxwell viscoelastic model has a larger misfit because it tends to overestimate the rate of deformation after about three years (Figure \ref{fig:RecordSectionMain}). Since our initial estimates of $\eta_\mathrm{eff}$ may be biased towards overestimating viscosities, we have also performed the slip inversion where we use uniformly lower viscosities in the crust and mantle. However, decreasing the viscosity only increases the misfit.  Although, the viscosities used here are consistent with the successful Maxwell viscoelastic models found by \citet{Rollins2015} and \citet{Spinler2015}, which had mantle viscosities on the order of $10^{18}$ Pa s and relatively higher lower crustal viscosities, we find that such a model is incapable of describing the entire postseismic time series.  \citet{Pollitz2001} similarly recognized this deficiency in a Maxwell rheology, which then motivated their exploration of a Burgers rheology upper mantle \citep{Pollitz2003}.

Instead of exploring a Burgers rheology mantle, which introduces two new parameters that need to be estimated, the transient viscosity $\eta_{K}$ and transient shear modulus $\mu_{K}$, we first consider a Zener rheology for the mantle, which only introduces one additional unknown parameter, $\mu_{K}$.  We assume that the lower crust still has a Maxwell rheology. The steady-state viscosity in the crust and the transient viscosity in the mantle are set equal to the inferred effective viscosities.  We then estimate the ratio of shear moduli $\frac{\mu_K}{\mu}$. We compute nine different sets of Green's functions, $f$ and $g$, where we assume values of $\frac{\mu_K}{\mu}$ ranging from 0 to 1. The former being a degenerate case where the Zener model reduces to the above Maxwell model.  We estimate coseismic slip and afterslip for each realization of $\frac{\mu_K}{\mu}$.  The shear moduli ratio that yields the best prediction to the observed postseismic displacements is found to be $0.375$ which produces a misfit of $\bar\chi^2=31.2$ (Figure \ref{fig:ShearModulusRatio}).  The improvement in the Zener model over the Maxwell model can be clearly seen in the fit to the far-field data (Figure \ref{fig:RecordSectionMain}). The Zener model does a significantly better job at explaining the transient rate of far-field deformation throughout the five years.

\begin{figure}
\includegraphics[scale=0.9]{Figures/RatioMisfit}
\centering 
\caption{Mean chi-squared value for the best fitting slip model as a function of the transient shear modulus relative to the elastic shear modulus in a Zener upper mantle. Large dot indicates our preferred ratio.}
\label{fig:ShearModulusRatio}
\end{figure}

@@

The regularized inference of coseismic slip and afterslip for our preferred Zener model is shown in Figure \ref{fig:FinalSlip}.  The inferred coseismic potency is $3.0\times10^{9}\ \mathrm{m}^3$, equivalent to a Mw=7.26 earthquake, and most of the slip is shallow and on the Sierra Cucapah fault segment.  The potency of five years of afterslip is $1.1\times10^{9}\ \mathrm{m}^3$. Most of the afterslip in our preferred model occurs within the first year after the earthquake and coincides with the location of our inferred coseismic slip. Inferred afterslip within the first year is accounting for the most rapid near-field transient deformation. After one year, afterslip is inferred to be deeper down on the Sierra Cucapah segment, which is describing much of the sustained near-field postseismic deformation.  We emphasize, that the GPS station closest to where we infer afterslip, P496, is still about 30 km away, which is too far for us to conclusively argue for sustained localized deformation rather than shallow distributed deformation.  The deep afterslip inferred after one year could potentially be describing deformation resulting from lower crustal flow. To test this, we have modified our preferred model by decreasing the lower crustal viscosity from $5.91\times10^{19}$ Pa s to $1\times10^{19}$ Pa s, which is still consistent with our viscosity inference from Section \ref{sec:InitialInversion}, and inverted for fault slip.  We find that a model with a weaker lower crust adequately describes the postseismic displacements without any afterslip after one year, while still requiring about the same amount of afterslip over the first year. We do believe that the early, shallow afterslip on the Sierra Cucapah segment is a robust feature in our preferred model, while we are not confident in our inference of later deep afterslip.

the postseismic displacements predicted by our preferred model are shown in Figures \ref{fig:NearField}, \ref{fig:FarField} and \ref{fig:RecordSectionMain}.  Overall, the trends in the near-field and far-field transient deformation are accurately described.  In particular, the trends in far-field deformation are much better described by our preferred Zener model than either an elastic model or a model with a Maxwell viscoelastic mantle (Figure \ref{fig:RecordSectionMain}).  There are a few areas where we have notable misfit.  Most of our misfit is for the near-field stations in the Imperial Valley, and we attribute this misfit to our relatively simple fault geometry, which does not account for potential fault slip in the Imperial Valley triggered by the El Mayor-Cucapah earthquake \citep{Wei2011a,Wei2015}. In particular, we are unable to model the sustained, rapid rate of deformation at station P496, which suggests that this station could be influenced by a more localized deformation mechanism than is considered in this study.  Additionally, we see systematic misfit in the later postseismic period west of the Landers and Hector Mine earthquakes, which may be the result of unmodeled postseismic deformation resulting from those earthquakes.  Lastly, there are clear discrepancies between the observed and predicted vertical deformation following the first year after the El Mayor Cucapah earthquake. We observe a broad uplift throughout Southern California, which is inconsistent with any postseismic model.

\begin{figure}
\includegraphics[scale=0.9]{Figures/FinalSlip}
\centering 
\caption{Inferred coseismic slip and afterslip when assuming a Maxwell rheology in the lower crust and a Zener rheology in the upper mantle.  The transient viscosity $\eta_K$ in the mantle and steady-state viscosity $\eta_M$ in the crust are set equal to the effective viscosities from Figure \ref{fig:EffectiveViscosity}. We set $\frac{\mu_K}{\mu}=0.375$ in the upper mantle.}
\label{fig:FinalSlip}
\end{figure}

@@

An additional deficiency with the Maxwell rheology is that it predicts a steady decay in the rate of postseismic deformation over time, which fails to describe the commonly observed rapid, early transience followed by a relatively steady rate of postseismic deformation.  One could explain the early transient postseismic deformation with fault creep and the later phase with relaxation in a Maxwell viscoelastic lower crust and upper mantle \citep[e.g][]{Hearn2009,Johnson2009}. However, postseismic deformation at distances greater than ${\sim}200$ km from the El Mayor-Cucapah epicenter can only be attributed to viscoelastic relaxation \citep{Freed2007a} and we have demonstrated that the far-field deformation cannot be explained with a Maxwell rheology (Figure \ref{fig:RecordSectionMain}).
